---
title: "Sustainability Analytics: Project Report Group 2"
subtitle: "Effect of Climate Change on Aletschgletscher and the village of Brig"
author:
  - name: "Vani Gupta"
    affiliation: "Hochschule Luzern"
    email: "vani.gupta@stud.hslu.ch"
  - name: "Ramona Kölliker"
    affiliation: "Hochschule Luzern"
    email: "ramona.kölliker@stud.hslu.ch"
  - name: "Valérie Lüthi"
    affiliation: "Hochschule Luzern"
    email: "valerie.lüthi@stud.hslu.ch"
  - name: "Alois Wagner"
    affiliation: "Hochschule Luzern"
    email: "alois.wagner@stud.hslu.ch"
date: "2025-09-05"
output:
  html_document:
    number_sections: true
    theme: lumen
    code_folding: hide
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    toc-title: Table of Content
  minidown::mini_document:
    result_folding: show
---

<u>**Loading R-Packages**</u>

```{r Loading_Packages, include=TRUE, warning=FALSE, message=FALSE, echo=TRUE, results='hide'}
#Loading needed packages
library(readr)
library(dplyr)
library(lubridate)
library(tseries)
library(zoo)
library(forecast)
```

<br><p style="font-size: 35px;"> Introduction </p><br>

This report presents the course project developed during the blockweek Sustainability Analytics FS25.



<u>**Loading Datasets**</u>

```{r loading_dataset, include=TRUE, warning=FALSE, message=FALSE, echo=TRUE, results='hide'}
# Reading in datasets
#df_data <- read.csv("xxxxx.csv")

# load Jungfraujoch data
df_jungfraujoch <- read.delim("../data/raw/weather_monthly_jungfraujoch.csv", header=TRUE, sep = ";")
head(df_jungfraujoch)
str(df_jungfraujoch)
```

# Data Preparation

<u>**Datapreparation of Jungfraujoch_Temperature_dataset**</u>
```{r jungfraujoch_temperature_datapreparation, include=TRUE, warning=FALSE, message=FALSE, echo=TRUE, results='hide'}
#code
df_jungfraujoch <- df_jungfraujoch %>%
  mutate(reference_timestamp = dmy_hm(reference_timestamp))
str(df_jungfraujoch)
```

<u>**Datapreparation of Brig_Rain dataset**</u>
```{r rain_brig_datapreparation, include=TRUE, warning=FALSE, message=FALSE, echo=TRUE, results='hide'}
#code
```


## H2

### H3 {#chapter-with-features}

<u>**----->"Header for the Code Block1<------**</u>
```{r first_subset_feature_exp, include=TRUE, warning=FALSE, message=FALSE, echo=TRUE}
#code
```
<u>**----->"Header for the Code Block1<------**</u>
```{r first_subset_feature_exp1, include=TRUE, warning=FALSE, message=FALSE, echo=TRUE}
#code
```

The [feature explanation](#Table_feature_exp) can be found in the appendix.






# Appendix

## Feature Explanation of the dataset {#Table_feature_exp}

This section provides further explanations for the dataset.

```{r data_dict, include=TRUE, warning=FALSE, message=FALSE, echo=TRUE}

```

[Back to Chapter 'H3'](#chapter-with-features)
