---
title: "Sustainability Analytics: Project Report Group 2"
subtitle: "Effect of Climate Change on Aletschgletscher and the village of Brig"
author:
  - name: "Vani Gupta"
    affiliation: "Hochschule Luzern"
    email: "vani.gupta@stud.hslu.ch"
  - name: "Ramona Kölliker"
    affiliation: "Hochschule Luzern"
    email: "ramona.kölliker@stud.hslu.ch"
  - name: "Valérie Lüthi"
    affiliation: "Hochschule Luzern"
    email: "valerie.lüthi@stud.hslu.ch"
  - name: "Alois Wagner"
    affiliation: "Hochschule Luzern"
    email: "alois.wagner@stud.hslu.ch"
date: "2025-09-05"
output:
  html_document:
    number_sections: true
    theme: lumen
    code_folding: hide
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    toc-title: Table of Content
  minidown::mini_document:
    result_folding: show
---

<u>**Loading R-Packages**</u>

```{r Loading_Packages, include=TRUE, warning=FALSE, message=FALSE, echo=TRUE, results='hide'}
#Loading needed packages
library(dplyr)
library(forecast)
library(ggplot2)
library(lubridate)
library(readr)
library(tseries)
library(zoo)
```

<br><p style="font-size: 35px;"> Introduction </p><br>

This report presents the course project developed during ........


# Data

<u>**Loading Datasets**</u>

```{r loading_dataset, include=TRUE, warning=FALSE, message=FALSE, echo=TRUE, results='hide'}
# Reading in datasets
#df_data <- read.csv("xxxxx.csv")

# load Jungfraujoch weather data
df_jungfraujoch <- read.delim("../data/raw/weather_monthly_jungfraujoch.csv", header=TRUE, sep = ";")
head(df_jungfraujoch)
str(df_jungfraujoch)
```
## Jungfraujoch Weather Data

The homogeneous data series from the Jungfraujoch climate station is available on demand via [MeteoSchweiz OpenData](https://www.meteoswiss.admin.ch/services-and-publications/service/open-data.html). The station is located 3571 meters above sea level and located in the Aletsch area. We were interested in the measurement of the homogeneous monthly mean temperature covering the time span from 01/01/1933 to 01/07/2025. The temperature in Celsius was measured 2 meters above ground.

# H1

<u>**Datapreparation of Brig_Rain dataset**</u>
```{r rain_brig_datapreparation, include=TRUE, warning=FALSE, message=FALSE, echo=TRUE, results='hide'}
#code
```
## H2

### H3 {#chapter-with-features}

<u>**----->"Header for the Code Block1<------**</u>
```{r first_subset_feature_exp, include=TRUE, warning=FALSE, message=FALSE, echo=TRUE}
#code
```
<u>**----->"Header for the Code Block1<------**</u>
```{r first_subset_feature_exp1, include=TRUE, warning=FALSE, message=FALSE, echo=TRUE}
#code
```

The [feature explanation](#Table_feature_exp) can be found in the appendix.






# Appendix

## Feature Explanation of the dataset {#Table_feature_exp}

This section provides further explanations for the dataset.

```{r data_dict, include=TRUE, warning=FALSE, message=FALSE, echo=TRUE}

```

[Back to Chapter 'H3'](#chapter-with-features)
