---
title: "Sustainability Analytics: Project Report Group 2"
subtitle: "Effect of Climate Change on Great Aletsch Glacier and the village of Brig"
author:
  - name: "Vani Gupta"
    affiliation: "Hochschule Luzern"
    email: "vani.gupta@stud.hslu.ch"
  - name: "Ramona Kölliker"
    affiliation: "Hochschule Luzern"
    email: "ramona.kölliker@stud.hslu.ch"
  - name: "Valérie Lüthi"
    affiliation: "Hochschule Luzern"
    email: "valerie.lüthi@stud.hslu.ch"
  - name: "Alois Wagner"
    affiliation: "Hochschule Luzern"
    email: "alois.wagner@stud.hslu.ch"
date: "2025-09-05"
output:
  html_document:
    number_sections: true
    theme: lumen
    code_folding: hide
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    toc-title: Table of Content
  minidown::mini_document:
    result_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<u>**Loading R-Packages**</u>
```{r Loading_Packages, include=TRUE, warning=FALSE, message=FALSE, echo=TRUE, results='hide'}
#Loading needed packages
library(tseries)  # for working with time series
library(ismev)    # for EVT gev.fit
library(tidyverse)# for data handling (includes ggplot2, dplyr, tidyr, readr, lubridate)
library(forecast)
library(evd)      # for calculating HQ values in EVT

```




# Introduction (Ramona)

<br><p style="font-size: 35px;"> Introduction </p><br>

Hinweis für Ramona: Falls wir für die Einleitung oder Abstrakt keine Auflistung im Inhaltsverzeichnis haben möchten, könnten wir das mit **'p style="font-size: 35px;"> Introduction </p'** anstelle von **## Introduction** lösen.


# Datapreparation/Cleaning

<u>**Loading Data**</u>
```{r loading_data, include= TRUE, echo=TRUE}
# Reading in data
raw_massa <- read.csv('../data/processed/massa_cleaned.csv', header=TRUE, sep=',', na.strings='-')
raw_rhone <- read.csv('../data/processed/rhone_cleaned.csv', header=TRUE, sep=',', na.strings='-')
raw_glacier <- read.csv('../data/processed/glacier_cleaned.csv', header=TRUE, sep=',', na.strings='-')
raw_glacier_monthly <- read.csv('../data/processed/glacier_monthly_processed.csv', header=TRUE, sep=',', na.strings='-')
raw_jungfraujoch <- read.csv('../data/processed/jungfraujoch_temperature.csv', header=TRUE, sep=',', na.strings='-')

```


## River Massa (Ramona)
## Glacier (Vani)


# Exploratory Data Analysis (Vani)


# Stationarity

## Temperature Jungfraujoch (Valérie)

## Rainfall Brig (Alois)


## River Massa
### Waterlevel (Ramona)

## River Rhone
### Waterlevel (Ramona)
### Watervolume (Alois)


# Modeling

## Crosscorrelation (Valérie)
## ARIMAX

## EVT (Alois)
### River Massa

### River Rhone
#### First data analysis rhone

<u>**EVT Rhone Data Preparation**</u>
```{r evt_rhone_preparation, echo=FALSE}
# Read data
df_rhone <- raw_rhone

## Time Series Rhone River - daily data
freq_daily <- 365.2422

ts_watervolume_rhone <- ts(df_rhone$discharge_vol_m3s, 
              start = c(year(min(df_rhone$date_yyyymmdd)),
                        yday(min(df_rhone$date_yyyymmdd))), 
              frequency = freq_daily)


# creating dataframe out of ts_watervolume_rhone 
#----------------------------
# defining start date
rhone_start_date <- as.Date("1981-01-01")
# length of time series
n <- length(ts_watervolume_rhone)
# creating date
rhone_dates <- seq.Date(from = rhone_start_date, by = "day", length.out = n)

df_rhone_prep <- data.frame(
  Date = rhone_dates,
  Watervolume = ts_watervolume_rhone
)

#----------------------------------------------------------------------
#----------------------------------------------------------------------
#First insights of rhone data / creating ts

#plotting rhone data / watervolume
plot(ts_watervolume_rhone, main = "Rhone daily discharge volume in m^3/s 1980 - 2020")
```

<u>**EVT Rhone Histogram**</u>
```{r EVT_histogram_rhone, include = TRUE, warning=FALSE, message=FALSE,}
# histogram
p_rhone <- ggplot(df_rhone_prep, aes(x=Watervolume)) + 
  geom_histogram() +
  ggtitle("Histogramm of rhone watervolume in m^3/s")
plot(p_rhone)
```

#### Plotting lm-trend line

<u>**EVT Rhone Data Preparation**</u>
```{r EVT_rhone, include = TRUE, warning=FALSE, message=FALSE,}
# defining trend lm
rhone_t_lm <- lm(Watervolume ~ Date, data = df_rhone_prep)

# defining trend lm with extrem values
threshold <- quantile(df_rhone_prep$Watervolume, 0.95)  # upper 5%-quantile
rhone_ex <- subset(df_rhone_prep, Watervolume > threshold)

rhone_ex_t_lm <- lm(Watervolume ~ Date, data = rhone_ex)

# storing lm coef in dataframe
trend_df <- data.frame(
  intercept = c(coef(rhone_t_lm)[1], coef(rhone_ex_t_lm)[1]),
  slope     = c(coef(rhone_t_lm)[2], coef(rhone_ex_t_lm)[2]),
  type      = c("Trend with all values", "Trend with extreme values")
)

# Plotting raw data with trend lines
ggplot(df_rhone_prep, aes(x = Date, y = Watervolume)) +
  geom_line(color = "black") +
  geom_abline(data = trend_df,
              aes(intercept = intercept, slope = slope, color = type),
              size = 1) +
  labs(title = "Watervolum with Trend lines",
       x = "Year", y = expression("m"^3*"/s"),
       color = "Trend type") +
  theme_minimal()
```




#### EVT Peaks over Threshold (POT)

<u>**EVT POT Rhone**</u>
```{r EVT_modeling_threshold_rhone, include = TRUE}
#EVT original data on upper quantil
threshold_rhone_evt <- quantile(df_rhone_prep$Watervolume, 0.998)
rhone_ex <- df_rhone_prep %>% filter(df_rhone_prep$Watervolume > threshold_rhone_evt)
rhone_ex_fit <- gev.fit(rhone_ex$Watervolume)
gev.diag(rhone_ex_fit) # -> Return Period is on a daily scale
rhone_ex_fit$mle
```

#### EVT Block Max Year 

<u>**EVT Block max year Rhone**</u>
```{r EVT_modeling_block_max_rhone, include = TRUE}
#-----------------
#EVT original data block max year

df_rhone_prep <- df_rhone_prep %>%
  mutate(Date = rhone_start_date + (Date - min(Date)))  # Time as day offset

rhone_max_year <- df_rhone_prep %>%
  mutate(YEAR = year(Date),
         MONTH = month(Date)) %>%
  group_by(YEAR) %>%
  summarize(Watervolume_MAX = max(Watervolume, na.rm = TRUE)) %>%
  ungroup()

rhone_max_year_fit <- gev.fit(rhone_max_year$Watervolume_MAX)
gev.diag(rhone_max_year_fit) # -> Return Period is on a yearly scale
rhone_max_year_fit$mle
```

### EVT HQ Values

```{r EVT_modeling_hq_rhone, include = TRUE}
# Outputting HQ Values
# -> -> Hochwasser-Quantil / Hochwasserabluss: flood quantil for a defined return interval
# -> HQ2 means the watervolume will be achieved/exceeded each 2 years

loc_rhone <- rhone_max_year_fit$mle[1]
scale_rhone <- rhone_max_year_fit$mle[2]
shape_rhone <- rhone_max_year_fit$mle[3]

# return periode in years
return_periods_rhone <- c(2, 10, 30, 100, 300)

# probability p = 1 - 1/T
p <- 1 - 1 / return_periods_rhone


# calculating HQ-Value
HQ_rhone <- qgev(p, loc=loc_rhone, scale=scale_rhone, shape=shape_rhone)

# output
hq_table_rhone <- data.frame(ReturnPeriod = return_periods_rhone, HQ = HQ_rhone)
hq_table_rhone
```

# Conclusion

# Outlook (Vani)


# Disclaimer for AI


# Appendix






























# Examples for formatting

<u>**Datapreparation of Brig_Rain dataset**</u>
```{r rain_brig_datapreparation, include=TRUE, warning=FALSE, message=FALSE, echo=TRUE, results='hide'}
#code
```
## H2

### H3 {#chapter-with-features}

<u>**----->"Header for the Code Block1<------**</u>
```{r first_subset_feature_exp, include=TRUE, warning=FALSE, message=FALSE, echo=TRUE}
#code
```
<u>**----->"Header for the Code Block1<------**</u>
```{r first_subset_feature_exp1, include=TRUE, warning=FALSE, message=FALSE, echo=TRUE}
#code
```

The [feature explanation](#Table_feature_exp) can be found in the appendix.






## Appendix

### Feature Explanation of the dataset {#Table_feature_exp}

This section provides further explanations for the dataset.

```{r data_dict, include=TRUE, warning=FALSE, message=FALSE, echo=TRUE}

```

[Back to Chapter 'H3'](#chapter-with-features)
