---
title: "modeling_extreme_event_massa_markdown"
date: "2025-09-08"
output:
  html_document:
    # css: RawData/CSS/styles.css
    number_sections: true
    theme: lumen
    code_folding: hide
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    toc-title: Table of Content
  minidown::mini_document:
    result_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r library, echo=FALSE, message=FALSE, warning=FALSE}
library(tseries)  # for working with time series
library(ismev)    # for EVT gev.fit
library(tidyverse)
library(forecast)
library(evd)      # for calculating HQ values
```

## First data analysis massa


```{r loading_data, echo=FALSE}
# Read data
df_massa <- read.csv('../data/processed/massa_cleaned.csv', header=TRUE, sep=',', na.strings='-')

#----------------------------------------------------------------------
#----------------------------------------------------------------------
#First insights of massa data / creating ts

# defining starting year and month
start_year  <- year(min(df_massa$date_yyyymmdd))
start_month <- month(min(df_massa$date_yyyymmdd))
start_day <- day(min(df_massa$date_yyyymmdd))

## Time Series Massa River - daily data
freq_daily <- 365.2422

ts_watervolume <- ts(df_massa$discharge_vol_m3s, 
              start = c(year(min(df_massa$date_yyyymmdd)),
                        yday(min(df_massa$date_yyyymmdd))), 
              frequency = freq_daily)

#plotting massa data / watervolume
plot(ts_watervolume, main = "Original Data")
```

## Making data stationary

### Daily data

#### decompose

```{r decompose, include = TRUE}
ts_watervolume_comp <- decompose(ts_watervolume)
plot(ts_watervolume_comp)

acf(na.omit(ts_watervolume_comp$random), main = "ACF of decompose residuals")
pacf(na.omit(ts_watervolume_comp$random), main = "PACF of decompose residuals")
plot(ts_watervolume_comp$random, main = "Residuals decompose")
adf.test(na.omit(ts_watervolume_comp$random))
```

#### Decompose + Diff lag = 365

```{r decompose_diff, include = TRUE}
ts_watervolume_comp_diff <- diff(ts_watervolume_comp$random, lag = 365)
acf(na.omit(ts_watervolume_comp_diff), main = "ACF of decompose + diff residuals")
pacf(na.omit(ts_watervolume_comp_diff), main = "PACF of decompose + diff residuals")
plot(ts_watervolume_comp_diff, main = "Residuals decompose + diff")
adf.test(na.omit(ts_watervolume_comp_diff))
```

#### stl approach

```{r stl, include = TRUE}
ts_watervolume_stl <- stl(ts_watervolume, s.window = 182.5, t.window=365.2422)
remainder_ts_watervolume_stl <- ts_watervolume_stl$time.series[,"remainder"]
acf(remainder_ts_watervolume_stl, main = "ACF of residuals stl")
pacf(remainder_ts_watervolume_stl, main = "PACF of residuals stl")
plot(remainder_ts_watervolume_stl, main = "Residuals stl")
adf.test(na.omit(remainder_ts_watervolume_stl))
```


#### mstl approach

```{r mstl, include = TRUE}
# mstl approach: with year and winter/summer season
ts_watervolume_msts <- msts(ts_watervolume, seasonal.periods = c(365.2422, 182.5))
ts_watervolume_mstl <- mstl(ts_watervolume_msts, s.window = 7, 
                            t.window = 350, robust = FALSE)

# plotting the mstl components underneath
#par(mfrow = c(5,1), mar = c(3,4,2,1))  # mar = Ränder anpassen

# extracting components
mstl_trend <- ts_watervolume_mstl[, "Trend"]
mstl_seasonal1 <- ts_watervolume_mstl[, "Seasonal365.24"]
mstl_seasonal2 <- ts_watervolume_mstl[, "Seasonal182.5"]
mstl_remainder <- ts_watervolume_mstl[, "Remainder"]

plot(ts_watervolume, type = "l", main = "Original", ylab = "")
plot(mstl_trend, type = "l", main = "Trend mstl", ylab = "")
plot(mstl_seasonal1, type = "l", main = "Seasonal 1 (Year) mstl", ylab = "")
plot(mstl_seasonal2, type = "l", main = "Seasonal 2 (Half-Year) mstl", ylab = "")
plot(mstl_remainder, type = "l", main = "Residuals mstl", ylab = "")

# creating dataframe with mstl objects
#----------------------------
# defining start date
start_date <- as.Date("1981-01-01")
# length of time series
n <- length(ts_watervolume_msts)
# creating date
dates <- seq.Date(from = start_date, by = "day", length.out = n)
#----------------------------

# extracting mstl components
mstl_df <- data.frame(
  Date = dates,
  Original = ts_watervolume_mstl[, "Data"],                  # original values
  Trend = ts_watervolume_mstl[, "Trend"],     # trend
  Seasonal1_Year = ts_watervolume_mstl[, "Seasonal365.24"],  # first season / year
  Seasonal2_Halfyear = ts_watervolume_mstl[, "Seasonal182.5"],  # second season / half year
  Remainder = ts_watervolume_mstl[, "Remainder"]   # residuals
)
```
### Monthly data



#### Monthly Average

```{r monthly_average, include = TRUE}
# calculation average monthly watervolume
df_massa_monthly <- df_massa %>%
  mutate(Year = year(date_yyyymmdd),
         Month = month(date_yyyymmdd)) %>%
  group_by(Year, Month) %>%
  summarise(
    discharge_monthly_mean = mean(discharge_vol_m3s, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(Date = as.Date(paste(Year, Month, "01", sep = "-")))
  
## Time Series Massa River - monthly data
ts_watervolume_monthly <- ts(df_massa_monthly$discharge_monthly_mean, 
              start = c(year(min(df_massa_monthly$Date)),
                        month(min(df_massa_monthly$Date))), 
              frequency = 12)


plot(ts_watervolume_monthly)
```

##### decompose approach

```{r decompose_monthly, include = TRUE}
ts_watervolume_monthly_compose <- decompose(ts_watervolume_monthly)
plot(ts_watervolume_monthly_compose)

acf(na.omit(ts_watervolume_monthly_compose$random), main = "ACF of residuals monthly average decompose")
pacf(na.omit(ts_watervolume_monthly_compose$random), main = "PACF of residuals monthly average decompose")
plot(ts_watervolume_monthly_compose$random, main = "Residuals monthly average decompose")
adf.test(na.omit(ts_watervolume_monthly_compose$random))
```

##### stl approach

```{r stl_monthly, include = TRUE}
  
ts_watervolume_monthly_stl <- stl(ts_watervolume_monthly, s.window = 24, t.window=365.2422)
remainder_ts_watervolume_monthly_stl <- ts_watervolume_monthly_stl$time.series[,"remainder"]
acf(remainder_ts_watervolume_monthly_stl, main = "ACF of residuals monthly average stl")
pacf(remainder_ts_watervolume_monthly_stl, main = "PACF of residuals monthly average stl")
plot(remainder_ts_watervolume_monthly_stl, main = "Residuals monthly average stl")
adf.test(na.omit(remainder_ts_watervolume_monthly_stl))
plot(ts_watervolume_monthly_stl)
```

```{r}
#Creating dataframe of the prepared data
massa_monthly_mean_stl <- data.frame(
  Date      = as.Date(paste(df_massa_monthly$Year, df_massa_monthly$Month, "01", sep = "-")),  # monthly date
  Original  = as.numeric(ts_watervolume_monthly),                 # original data
  Trend     = ts_watervolume_monthly_stl$time.series[,"trend"],         # trend component
  Seasonal  = ts_watervolume_monthly_stl$time.series[,"seasonal"],      # seasonal component
  Remainder = ts_watervolume_monthly_stl$time.series[,"remainder"]      # remainder
)

head(massa_monthly_mean_stl)  # for checking

# # CSV export
# write.csv(
#   massa_monthly_mean_stl, 
#   "stationary_monthly_mean_watervolume_massa.csv", 
#   row.names = FALSE
# )
# 
# # RDS export
# saveRDS(
#   massa_monthly_mean_stl, 
#   "stationary_monthly_mean_watervolume_massa.rds"
# )
```


#### Monthly Max

```{r monthly_max, include = TRUE}
# calculation max monthly watervolume
df_massa_monthly_max <- df_massa %>%
  mutate(Year = year(date_yyyymmdd),
         Month = month(date_yyyymmdd)) %>%
  group_by(Year, Month) %>%
  summarise(
    discharge_monthly_max = max(discharge_vol_m3s, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(Date = as.Date(paste(Year, Month, "01", sep = "-")))
  
## Time Series Massa River - monthly data
ts_watervolume_monthly_max <- ts(df_massa_monthly_max$discharge_monthly_max, 
              start = c(year(min(df_massa_monthly_max$Date)),
                        month(min(df_massa_monthly_max$Date))), 
              frequency = 12)


plot(ts_watervolume_monthly_max)
```

##### decompose approach

```{r decompose_monthly_max, include = TRUE}
ts_watervolume_monthly_max_compose <- decompose(ts_watervolume_monthly_max)
plot(ts_watervolume_monthly_max_compose)

acf(na.omit(ts_watervolume_monthly_max_compose$random), main = "ACF of residuals monthly max decompose")
pacf(na.omit(ts_watervolume_monthly_max_compose$random), main = "PACF of residuals monthly max decompose")
plot(ts_watervolume_monthly_max_compose$random, main = "Residuals monthly max decompose")
adf.test(na.omit(ts_watervolume_monthly_max_compose$random))
```

##### stl approach

```{r stl_monthly_max, include = TRUE}
  
ts_watervolume_monthly_max_stl <- stl(ts_watervolume_monthly_max, s.window = 24, t.window=530)
remainder_ts_watervolume_monthly_max_stl <- ts_watervolume_monthly_max_stl$time.series[,"remainder"]
acf(remainder_ts_watervolume_monthly_max_stl, main = "ACF of residuals monthly max stl")
pacf(remainder_ts_watervolume_monthly_max_stl, main = "PACF of residuals monthly max stl")
plot(remainder_ts_watervolume_monthly_max_stl, main = "Residuals monthly max stl")
adf.test(na.omit(remainder_ts_watervolume_monthly_max_stl))
plot(ts_watervolume_monthly_max_stl)
```

## EVT modelling

### plotting lm-trend line
```{r EVT, include = TRUE}
# Plotting extracted mstl trend and lm-trend/abline
# preparing data
trend_df <- data.frame(
  Time = as.numeric(mstl_df$Date),
  Trend = mstl_df$Trend
)

# linear regression on trend
trend_t_lm <- lm(Trend ~ Time, data = trend_df)

# Plot
plot(trend_df$Time, trend_df$Trend, type="l", main="mstl trend with regression line", xlab="Year", ylab="Trend")
abline(trend_t_lm, col="red", lwd=2)

#----------------------------
# Plotting original data with lm-trend/abline
orig_df <- data.frame(
  Time = as.numeric(mstl_df$Date),
  Watervolume = mstl_df$Original
)

orig_t_lm <- lm(Watervolume ~ Time, data = orig_df)

plot(orig_df$Time, orig_df$Watervolume, type="l", main="Original with regression trend-line", xlab="Year", ylab="m^3/s")
abline(orig_t_lm, col="blue", lwd=2)

# Plotting original data with normal trend/abline and extrem value lm-trend/abline
threshold <- quantile(orig_df$Watervolume, 0.95)  # upper 5%-quantile
orig_ex <- subset(orig_df, Watervolume > threshold)

orig_ex_t_lm <- lm(Watervolume ~ Time, data = orig_ex)

# Plot
plot(orig_df$Time, orig_df$Watervolume, type="l", main="Original + extrem trend", xlab="Year", ylab="m^3/s")
abline(orig_t_lm, col="blue", lwd=2)
abline(orig_ex_t_lm, col="green", lwd=2)


# Comparing both plots on same scale

# setting plot settings
par(mfrow = c(2,1), mar = c(5,4,3,1))

# common scaling of y-axis for plotting lm-trend
ymin <- min(c(mstl_df$Trend, mstl_df$Original), na.rm = TRUE)
ymax <- max(c(mstl_df$Trend, mstl_df$Original), na.rm = TRUE)

# mstl trend-component
plot(mstl_df$Date, mstl_df$Trend, type="l", ylim=c(ymin, ymax),
     main="mstl trend-Component", xlab="Year", ylab="m^3/s")
abline(lm(Trend ~ as.numeric(Date), data=mstl_df), col="red", lwd=2)

# original data
plot(mstl_df$Date, mstl_df$Original, type="l", ylim=c(ymin, ymax),
     main="Original Data", xlab="Year", ylab="m^3/s")
abline(lm(Original ~ as.numeric(Date), data=mstl_df), col="blue", lwd=2)

# resetting plot settings
par(mfrow = c(1,1), mar = c(5,4,3,1))
```

```{r EVT_histogram_massa, include = TRUE}
# histogram
p_massa <- ggplot(orig_df, aes(x=Watervolume)) + 
  geom_histogram() +
  ggtitle("Histogramm of massa watervolume in m^3/s")
plot(p_massa)
```

### EVT threshold quantile > 0.998

```{r EVT_modeling_threshold, include = TRUE}
#EVT original data on upper quantil
threshold2 <- quantile(orig_df$Watervolume, 0.998)
original_ex <- orig_df %>% filter(orig_df$Watervolume > threshold2)
original_ex_fit <- gev.fit(original_ex$Watervolume)
gev.diag(original_ex_fit) # -> Return Period is on a daily scale
original_ex_fit$mle
```

### EVT block max year 

```{r EVT_modeling_block_max, include = TRUE}
#-----------------
#EVT original data block max year

orig_df <- orig_df %>%
  mutate(Date = start_date + (Time - min(Time)))  # Time as day offset

orig_df_max_year <- orig_df %>%
  mutate(YEAR = year(Date),
         MONTH = month(Date)) %>%
  group_by(YEAR) %>%
  summarize(Watervolume_MAX = max(Watervolume, na.rm = TRUE)) %>%
  ungroup()

orig_df_max_year_fit <- gev.fit(orig_df_max_year$Watervolume_MAX)
gev.diag(orig_df_max_year_fit) # -> Return Period is on a yearly scale
orig_df_max_year_fit$mle
```

### EVT HQ Values

```{r EVT_modeling_hq, include = TRUE}
# Outputting HQ Values
# -> -> Hochwasser-Quantil / Hochwasserabluss: flood quantil for a defined return interval
# -> HQ2 means the watervolume will be achieved/exceeded each 2 years

loc <- orig_df_max_year_fit$mle[1]
scale <- orig_df_max_year_fit$mle[2]
shape <- orig_df_max_year_fit$mle[3]

# return periode in years
return_periods <- c(2, 10, 30, 100, 300)

# probability p = 1 - 1/T
p <- 1 - 1 / return_periods


# calculating HQ-Value
HQ <- qgev(p, loc=loc, scale=scale, shape=shape)

# output
hq_table <- data.frame(ReturnPeriod = return_periods, HQ = HQ)
hq_table
```



## First data analysis rhone

```{r loading_data2, echo=FALSE}
# Read data
df_rhone <- read.csv('../data/processed/rhone_cleaned.csv', header=TRUE, sep=',', na.strings='-')

#----------------------------------------------------------------------
#----------------------------------------------------------------------
#First insights of rhone data / creating ts

# defining starting year and month
start_year2  <- year(min(df_rhone$date_yyyymmdd))
start_month2 <- month(min(df_rhone$date_yyyymmdd))
start_day2 <- day(min(df_rhone$date_yyyymmdd))

## Time Series Massa River - daily data
freq_daily <- 365.2422

ts_watervolume_rhone <- ts(df_rhone$discharge_vol_m3s, 
              start = c(year(min(df_rhone$date_yyyymmdd)),
                        yday(min(df_rhone$date_yyyymmdd))), 
              frequency = freq_daily)

#plotting rhone data / watervolume
plot(ts_watervolume_rhone, main = "Original Data")
```


### plotting lm-trend line
```{r EVT_rhone, include = TRUE}

# creating dataframe with mstl objects
#----------------------------
# defining start date
start_date <- as.Date("1981-01-01")
# length of time series
n <- length(ts_watervolume_rhone)
# creating date
dates <- seq.Date(from = start_date, by = "day", length.out = n)

rhone_df <- data.frame(
  Date = dates,
  Watervolume = ts_watervolume_rhone
)

#-plot
orig_t_lm <- lm(Watervolume ~ Date, data = rhone_df)

plot(rhone_df$Date, rhone_df$Watervolume, type="l", main="Watervolume with regression trend-line", xlab="Year", ylab="m^3/s")
abline(orig_t_lm, col="blue", lwd=2)

# Plotting original data with normal trend/abline and extrem value lm-trend/abline
threshold <- quantile(rhone_df$Watervolume, 0.95)  # upper 5%-quantile
orig_ex <- subset(rhone_df, Watervolume > threshold)

orig_ex_t_lm <- lm(Watervolume ~ Date, data = orig_ex)

# Plot
plot(rhone_df$Date, rhone_df$Watervolume, type="l", main="Original + extrem trend", xlab="Year", ylab="m^3/s")
abline(orig_t_lm, col="blue", lwd=2)
abline(orig_ex_t_lm, col="green", lwd=2)

```

```{r EVT_histogram_rhone, include = TRUE}
# histogram
p_rhone <- ggplot(rhone_df, aes(x=Watervolume)) + 
  geom_histogram() +
  ggtitle("Histogramm of rhone watervolume in m^3/s")
plot(p_rhone)
```

### EVT threshold quantile > 0.998

```{r EVT_modeling_threshold_rhone, include = TRUE}
#EVT original data on upper quantil
threshold3 <- quantile(rhone_df$Watervolume, 0.998)
rhone_ex3 <- rhone_df %>% filter(rhone_df$Watervolume > threshold3)
rhone_ex_fit3 <- gev.fit(rhone_ex3$Watervolume)
gev.diag(rhone_ex_fit3) # -> Return Period is on a daily scale
rhone_ex_fit3$mle
```

### EVT block max year 

```{r EVT_modeling_block_max_rhone, include = TRUE}
#-----------------
#EVT original data block max year

rhone_df <- rhone_df %>%
  mutate(Date = start_date + (Date - min(Date)))  # Time as day offset

rhone_df_max_year <- rhone_df %>%
  mutate(YEAR = year(Date),
         MONTH = month(Date)) %>%
  group_by(YEAR) %>%
  summarize(Watervolume_MAX = max(Watervolume, na.rm = TRUE)) %>%
  ungroup()

rhone_df_max_year_fit <- gev.fit(rhone_df_max_year$Watervolume_MAX)
gev.diag(rhone_df_max_year_fit) # -> Return Period is on a yearly scale
rhone_df_max_year_fit$mle
```

### EVT HQ Values

```{r EVT_modeling_hq_rhone, include = TRUE}
# Outputting HQ Values
# -> -> Hochwasser-Quantil / Hochwasserabluss: flood quantil for a defined return interval
# -> HQ2 means the watervolume will be achieved/exceeded each 2 years

loc2 <- rhone_df_max_year_fit$mle[1]
scale2 <- rhone_df_max_year_fit$mle[2]
shape2 <- rhone_df_max_year_fit$mle[3]

# return periode in years
return_periods <- c(2, 10, 30, 100, 300)

# probability p = 1 - 1/T
p <- 1 - 1 / return_periods


# calculating HQ-Value
HQ <- qgev(p, loc=loc2, scale=scale2, shape=shape2)

# output
hq_table2 <- data.frame(ReturnPeriod = return_periods, HQ = HQ)
hq_table2
```